The Analogue Method for Precipitation Forecasting: Finding Better Analogue Situations at a Sub-Daily Time Step

### 2015-12-18 Moving temporal window - HESS/template.tex
Personal corrections.

\begin{abstract}
The Analogue Method (AM) aims at forecasting local weather variables (predictands), such as the precipitation, by means of a statistical relationship with predictors at a synoptic scale. The analogy is assessed in the first place on the geopotential field by mean of a comparison of the gradients, in order to sample the days with a similar atmospheric circulation.

The search for candidates situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day. The constraint being the use of daily time series, due to the length of available archives they provide. However, it is unlikely that the best analogues happen at the very same hour, but may occur at different time of the day. In order to assess the potential of finding better analogues at a different times, a sliding time window (STW) has been introduced.

The STW resulted in a better analogy of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analogue dates. The improvement was found to grow with the analogue ranks due to an accumulation of more similar situations in the predicted distribution. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

A seasonal effect has also been identified, with larger improvements in winter that in summer, supposedly due to the stronger effect of the diurnal cycle in summer, which favors predictors at the same hours.

@@

Attempts to reconstruct longer precipitation series of running 24~h-totals by means of simple methods failed. It empathized the need to use time series with an accurate chronology. These should be available in a near future, either by means of growing observed time series, or by the establishment of precipitation reanalysis through regional modeling. Then, the use of a STW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

The Analogue Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects, and can be called analogues \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The versions that will be considered here are often used as references or benchmarks for various improvements. The first implementation relies on an analogy of the atmospheric circulation, and the other adds a second level of analogy on moisture variables \citep{Obled2002, Bontron2005, Marty2012}.

The first method is based on the analogy of the synoptic circulation (Table \ref{table:method_2Z}). When searching analogue situations for a target day, it starts by a preselection step, which is a selection of the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of a target date with every day of the preselection is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window. The considered pressure levels are 500~hPa (Z500) and 1000~hPa (Z1000) \citep{Bontron2005}. The $N_{1}$ dates, which is a parameter to calibrate, with the lowest values of S1 are considered as analogues to the target day. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

@@

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique was used in order to calibrate the parameters in the present study. The score that is most often used to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent to a finer resolution. Therefore, the analogues situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations do not occur systematically at the same time of the day, and that better candidates can be found at different times. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes on a respectable archive length was used (section \ref{sec:data}). It was then aggregated in the form of 24~h totals, but at a 6-hourly time step, by means of a moving average (section \ref{sec:method}).

@@

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[study area of][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no aggregation or catchment rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle, which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations.

@@

In order to assess the benefit of searching analogue situations at a sub-daily time step, an appropriate time series is required. On the basis of high resolution time series (section \ref{sec:data}), 24~h totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving average.

@@

Logic suggests a priori that only one situation (the best) is selected per candidate day. An exclusion of the same dates was tested in two different ways: (1)
by excluding the worse analogues in a period of $\pm$ 24~h, and (2) by excluding the worse analogues with the same date. After several tests, it turned out that no restrictions led to better skills, and this on a significant and recurring basis. Thus, no constraint was further used.

@@

As sub-daily precipitation time series are usually available on a shorter period than traditional daily time steps, the first assessment consists in assessing the loss of performance resulting from a reduction a 47~years archive (1961 to 2008) to 25~years (1982 to 2007). This change is assessed with the original method, without STW.

Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated using AtmoSwing \citep{Horton2016} on the original archive (Tables \ref{table:method_2Z} and \ref{table:method_2Z-2MI}) and will be used thereafter.

@@

\section{Influence on the analogy criteria}

@@

The median of the S1 values reduction (Figure \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria has been improved for all analogue ranks of the distribution end. All other stations have a similar improvement of the S1 criteria, both in terms of distribution shape and amplitude.

@@

\section{Impact on performance scores}

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better forecast skills. Therefore, the impact of the introduction of the STW, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see section \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

The increasing positive trend of skill improvements regarding the precipitation threshold is significant and shows that the prediction of higher precipitation totals is further improved. Thus both the analogy criteria and the performance scores are improved to a greater extent for heavier precipitation events. On the contrary, the non-rainy days and small accumulations are not improved.

@@

The values of the CRPSS scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the STW, the performance loss related to the reduction of the archive length is retrieved. In our case, this improvement is approximately a doubling of the archive size.

@@

\section{Application attempts on the full archive}

The improvement provided by the STW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving averages of 24~h-totals, different disaggregation approaches of the daily time series were assessed and are presented in the following sections.

@@

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5\textdegree\~-~7.5\textdegree\ longitude and 45\textdegree\~-~47.5\textdegree\ latitude), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

@@

These attempts to transpose the STW on the total archive highlight the importance of the actual chronology of rainfall. The STW is profitable, provided that the precipitation series are close to the observed one. Without precipitation series with an accurate sub-daily chronology, the introduction of a STW is not appropriate.

@@

The parameters were then calibrated again, using the STW. Some parameters changed, the main difference being the number of analogues that was found optimal, which systematically and significantly increased compared to the original set. A wider selection of analogue situations, containing those whose rank decreased, seemed profitable for the prediction performance. The number of good analogues is thus globally increased. This change seems to benefit mainly the prediction of days with small precipitation totals.

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analogue situations identified by the STW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed.

>>>>

\begin{abstract}
The Analogue Method (AM) aims at forecasting local weather variables (predictands), such as precipitations, by means of a statistical relationship with predictors at a synoptic scale. The analogy is generally assessed in the first place on the geopotential field by mean of a comparison of the gradients, in order to sample the days with a similar atmospheric circulation.

The search for candidate situations, for a given target day, is usually undertaken by comparing the state of the atmosphere at fixed hours of the day. The constraint being the use of daily time series, due to the length of available archives they provide. However, it is unlikely that the best analogy happens at the very same hour, but it may occur at a different time of the day. In order to assess the potential of finding better analogues at a different hour, a sliding time window (STW) has been introduced.

The STW resulted in a better analogy of the atmospheric circulation, with improved values of the analogy criteria on the whole distribution of analogue dates. The improvement was found to grow with the analogue ranks due to an accumulation of more similar situations in selection. Moreover, the improvement is even more important for days with heavy precipitation events, which are generally related to more dynamic atmospheric situations, where timing is more specific.

A seasonal effect has also been identified, with larger improvements in winter than in summer, supposedly due to the stronger effect of the diurnal cycle in summer, which favors predictors at the same hour.

@@

Attempts to reconstruct longer precipitation series of running 24~h-totals by means of simple methods failed. It empathized the need to use time series with an accurate chronology. These should be available in a near future, either by means of growing observed archives, or by the establishment of precipitation reanalysis through regional modeling. Then, the use of a STW in the AM should be considered for any application, especially when the prediction quality of extreme events is important.

@@

The Analogue Method (AM) is based on the principle that two relatively similar synoptic situations may produce relatively similar local effects \citep{Lorenz1956, Lorenz1969}. Multiple variations of the methods exist and listings can be found in \citet{Horton2016} or \citet{BenDaoud2015}. The versions that will be considered here are often used as references or benchmarks for various improvements.

The first method is based on the analogy of the atmospheric circulation \citep[Table \ref{table:method_2Z},][]{Obled2002, Bontron2005, Marty2012}. When searching analogue situations for a target day, it starts by a preselection step, which is a selection of the 4 months centered around the target date for every year of the archive, in order to cope with seasonal effects. Then, the similarity of the atmospheric circulation of the target date with every day of the preselection is assessed by processing the S1 criteria \citep[Eq.\ \ref{eq:S1}, ][]{Teweles1954, Drosdowsky2003}, which is a comparison of gradients, over a certain spatial window. The considered pressure levels are 500~hPa (Z500) and 1000~hPa (Z1000). The $N_{1}$ dates with the lowest values of S1 are considered as analogues to the target day, $N_{1}$ being a parameter to calibrate. Then, the daily observed precipitation amount of the corresponding dates provide the empirical conditional distribution considered as the probabilistic forecast for the target day. Following the nomenclature proposed by \citet{Horton2016}, this method will be named 2Z.

@@

To calibrate the method, the commonly used technique is a semi-automatic sequential procedure elaborated by \citet{Bontron2004}. AtmoSwing \citep{Horton2016}, which implements this technique, was used in order to calibrate the parameters in the present study. The score that is most often considered to assess an AM performance is the CRPS \citep[Continuous Ranked Probability Score,][]{Brown1974, Matheson1976, Hersbach2000}. It allows evaluating the predicted cumulative distribution functions $F(y)$, for example of the precipitation values $y$ from analogue situations, compared to the observed value $y^{0}$. The better the forecast, the smaller the score. The mean CRPS of a forecast series of length $n$ can be written:

@@

The AM is usually implemented with a daily time step, due to the availability of long precipitation archives that have no equivalent to a finer resolution. Therefore, the analogue situations are assessed on the basis of a daily time step, by comparing predictors at fixed hours of the day, otherwise one would not know what precipitation values to assign to them. However, it can be expected that the analogy of the synoptic situations do not occur systematically at the same time of the day, and that better candidates can be found at a different hour. On this assumption, a sliding time window (STW) was introduced to allow searching for candidates at different hours of the day.

Previous tests showed the benefit, in terms of analogy criteria values, of searching for analogue synoptic situations at a finer time step, but without assessing the impact on the forecast skills \citep{Finet2008}. In order to assess this potential improvement of the prediction, precipitation data with a resolution of 10 minutes, on a respectable archive length, was used (section \ref{sec:data}). It was then aggregated in the form of 24~h-totals, but at a 6-hourly time step, by means of a moving average (section \ref{sec:method}).

@@

The changes due to the introduction of a STW was assessed first in regards to the analogy criteria improvement (section \ref{sec:influence_criteria}), and then in terms of precipitation prediction skill (section \ref{sec:influence_scores}). Finally, attempts were made to reconstruct longer 24~h-total precipitations series (section \ref{sec:attempts}).

@@

The study area is the upper Rh\^{o}ne catchment in Switzerland \cite[see also][]{Horton2012a}. Due to the low density of weather stations with high temporal resolution and long archives, no aggregated rainfall was processed. The time series (on the period 1982-2007) come from 6 automatic weather stations, namely Ulrichen, Zermatt, Visp, Montana, Sion and Aigle (Figure \ref{fig:map}), which are subject to various meteorological influences \citep{Horton2012}. The results will thereafter be presented arbitrarily for the Ulrichen station, but they are equivalent for all stations.

\begin{figure}[htb]
	\begin{center}
		\includegraphics[width=8.3cm]{figures/figure_map.pdf}
	\end{center}
	\caption{Position of the 6 weather stations of interest (Ulrichen, Zermatt, Visp, Montana, Sion and Aigle) in the study area.}
	\label{fig:map}
\end{figure}

@@

In order to assess the benefit of searching analogue situations at a sub-daily time step, an appropriate precipitation series is required. On the basis of high resolution time series (section \ref{sec:data}), 24~h-totals were processed, but at a 6-hourly time step (temporal resolution of the reanalysis dataset), by means of a moving 24~h-total.

@@

Logic suggests \textit{a priori} that only one situation (the best) is selected per candidate day. An exclusion of the same dates was tested in two different ways: (1)
by excluding the worse analogues in a period of $\pm$ 24~h, and (2) by excluding the worse analogues with the same date. After several tests, it turned out that no restrictions led to better skills, and this on a significant and recurring basis. Thus, no constraint was further used.

@@

As sub-daily precipitation time series are usually available on a shorter period than traditional daily time steps, the first assessment consists in assessing the loss of performance resulting from a reduction of a 47~years archive (1961 to 2008) to 25~years (1982 to 2007). This change is assessed with the original method, without STW.

Both 2Z (Table \ref{table:method_2Z}) and 2Z-2MI (Table \ref{table:method_2Z-2MI}) methods were considered. The AM parameters were calibrated using AtmoSwing \citep{Horton2016} on the original archive (Tables \ref{table:params_2Z} and \ref{table:params_2Z-2MI}) and will be used thereafter.

@@

\section{Influence on the analogy criteria}
\label{sec:influence_criteria}

@@

The median of the S1 values reduction (Figure \ref{fig:changes_S1} bottom) starts approximately at 5~\% for the first analogue and reaches more than 10~\% for the last one. This increasing trend with the analogue rank can be explained by the accumulation of better analogues in the distribution. The minimum improvement starts from 0 and reaches 2-3\%, meaning that the criteria has been improved on most analogues for every day of the calibration period. All other stations have a similar improvement of the S1 criteria, both in terms of distribution shape and amplitude.

@@

\section{Impact on performance scores}
\label{sec:influence_scores}

A systematic improvement of the S1 values was previously observed. However, finding better analogue situations does not obligatory imply better skills to predict the precipitation. Therefore, the impact of the STW introduction, and thus the selection of other analogue dates, has to be assessed on the performance scores. In order to perform this assessment, the 24~h-totals (moving average) at a 6-hourly time step were used (see section \ref{sec:data}). The target dates remain unchanged, since the original time slot (6~h~UTC - 6~h~UTC the next day) is kept, and thus the performance scores can be directly compared with the former ones.

@@

The increasing positive trend of skill improvements regarding the precipitation threshold is significant and shows that the prediction of higher precipitation totals is further improved. Thus, both the analogy criteria and the performance scores are improved to a greater extent for heavier precipitation events. On the contrary, the non-rainy days and small accumulations are not improved.

@@

The values of the CRPSS scores for both methods (Table \ref{table:CRPSS_recalibration}) have significantly increased after recalibration. With the introduction of the STW, the performance loss related to the reduction of the archive length is retrieved.

@@

\section{Application attempts on the full archive}
\label{sec:attempts}

The improvement provided by the STW is interesting, mainly for heavy precipitation events, and thus it would be profitable to be able to apply it to the complete archive. Unfortunately, there is no long precipitation time series available with a sub-daily time step. In order to reconstruct a longer archive of moving 24~h-totals, different disaggregation approaches of the daily time series were assessed and are presented in the following sections.

@@

The first step consisted in determining which variables from the precipitable water or the relative humidity (at 1000~hPa, 925~hPa or 850~hPa) produces the precipitation chronology that is the most correlated (on non-zero values) with the accurate 6-hourly precipitation time series on the period 1982-2007. The location of the most predictive point from the reanalysis grid had to be assessed (closest points: 5\textdegree\ and 7.5\textdegree\ longitude and 45\textdegree\ and 47.5\textdegree\ latitude), as well an eventual time lapse between both series, due to the significant distance separating the weather stations and the reanalysis grid point. The most relevant point will probably not be the same for all stations, since one can expect to find the optimum in the direction of the main moisture inflows for the station. This assessment must be carried out for each station.

@@

These attempts to transpose the STW on the total archive highlight the importance of the actual rainfall chronology. The STW is profitable, provided that the precipitation series are close to the observed one. Without precipitation series with an accurate sub-daily chronology, the introduction of a STW is not appropriate.

@@

The parameters were then calibrated again, using the STW. Some parameters changed, the main difference being the number of analogues, which systematically and significantly increased compared to the original set. A wider selection of analogue situations, containing those whose rank decreased, seemed profitable for the prediction performance. The number of good analogues is thus globally increased. This change seems to benefit mainly the prediction of days with small precipitation totals.

The importance of the quality of the precipitation archive was also demonstrated, as too simplistic reconstructions of 6-hourly time series lead to a loss of all the improvement brought by better analogue situations identified by the STW. The precipitation prediction is improved only when the precipitation chronology is close to the accurate one. Attempts to reconstruct longer time series based on a simplistic proportional distribution or using meteorological variables from the NCEP reanalysis 1 dataset as proxy did not succeed. Other reanalysis datasets could eventually preform better.
